Learning how to act: making good decisions with machine learning

### 2017-09-16 thesis/thesis_map.tex
adds bit about bandits and RL

Remarkably for models developed relatively independently in fields with very different approaches and problems, the models we have discussed can be nicely unified for interventional queries (those that can be expressed with the do-notation). If the network for a structural equation model is acyclic, that is if starting from any node and following edges in the direction of the arrows you cannot return to the starting point, then it implies a recursive factorisation of the joint distribution over its variables. In other words, the network is a causal Bayesian network. All of the results that apply to causal Bayesian networks also apply to acyclic structural equation models.  Taking an action that sets a variable to a specific value equates to replacing the equation for that variable with a constant. This corresponds to dropping a term in the factorisation and the truncated product formula (equation \ref{eq:truncatedproduct}). Thus, the interventional query $P(Y|do(X))$ is identical in these two frameworks. We can also connect this to counterfactuals via:

@@

The estimator in equation \ref{eqn:importance_weighted_erm} is an example of a doubly robust estimator \citep{Robins1994a,Kang2007}. Doubly robust methods are asymptotically unbiased so long as either the regression model $h$ or propensity score $e$ are correctly specified \citep{robins1995semiparametric}.

@@

\begin{figure}[ht]
\begin{subfigure}[t]{0.99\textwidth}
     \caption{No prior on $\beta$, the posterior on $w_{yx}$ is centered around its true value}
    \includegraphics[scale=1]{prior_no_prior}
\end{subfigure}
\begin{subfigure}[t]{0.99\textwidth}
     \caption{Prior on $\beta$ centered around $w_{yz}$, the causal effect of $Z$ on $Y$, $Prior(\beta)= N(w_{yz},\sigma = 0.5)$. The posterior on $w_{yx}$ is biased away from its true value (as is the posterior on $\beta$ but this is not the key quantity of interest).}
    \includegraphics[scale=1]{prior_bad_prior.pdf}
\end{subfigure}
\caption{An example demonstrating how selecting a prior for a nuisance parameter centered around the causal effect of that parameter on the outcome, rather than its association with the outcome, can bias estimates for actual parameter of interest. We sample $n=1000$ data points from the joint distribution defined by $P(U)P(Z|U)P(X|Z)P(Y|U,Z,X) \sim N(0,1)N(2U,0.09)N(0.5Z,1)N(3U -Z + 0.5X , 0.25)$ and fit the model $Y \sim N(w_yx X+\beta Z,\epsilon)$ in Stan and plot the posterior distributions over $w_yx$ and $\beta$. Figure (a) shows the results with no prior (equivalently an improper prior) for both $w_yx$ and $\beta$. Figure (b) shows the results when we place a Gaussian prior centered around the $w_yz$ on the distribution for $\beta$.}\label{fig:careful-prior-example}
\end{figure}

@@

>>>>

Remarkably for models developed relatively independently in fields with very different approaches and problems, the models we have discussed can be nicely unified for interventional queries (those that can be expressed with the do-notation) \citep{Pearl2000}. If the network for a structural equation model is acyclic, that is if starting from any node and following edges in the direction of the arrows you cannot return to the starting point, then it implies a recursive factorisation of the joint distribution over its variables. In other words, the network is a causal Bayesian network. All of the results that apply to causal Bayesian networks also apply to acyclic structural equation models.  Taking an action that sets a variable to a specific value equates to replacing the equation for that variable with a constant. This corresponds to dropping a term in the factorisation and the truncated product formula (equation \ref{eq:truncatedproduct}). Thus, the interventional query $P(Y|do(X))$ is identical in these two frameworks. We can also connect this to counterfactuals via:

@@

The estimator in equation \ref{eqn:importance_weighted_erm} is an example of a doubly robust estimator \c%fitep{Robins1994a,Kang2007}. Doubly robust methods are asymptotically unbiased so long as either the regression model $h$ or propensity score $e$ are correctly specified \citep{robins1995semiparametric}.

@@

\begin{figure}[ht]
\begin{subfigure}[t]{0.99\textwidth}
     \caption{No prior on $\beta$, the posterior on $w_{yx}$ is centered around its true value}
    \includegraphics[scale=1]{prior_no_prior}
\end{subfigure}
\begin{subfigure}[t]{0.99\textwidth}
     \caption{Prior on $\beta$ centered around $w_{yz}$, the causal effect of $Z$ on $Y$, $Prior(\beta)= N(w_{yz},\sigma = 0.5)$. The posterior on $w_{yx}$ is biased away from its true value (as is the posterior on $\beta$ but this is not the key quantity of interest).}
    \includegraphics[scale=1]{prior_bad_prior.pdf}
\end{subfigure}
\caption{An example demonstrating how selecting a prior for a nuisance parameter centered around the causal effect of that parameter on the outcome, rather than its association with the outcome, can bias estimates for actual parameter of interest. We sample $n=1000$ data points from the joint distribution defined by $P(U)P(Z|U)P(X|Z)P(Y|U,Z,X) \sim N(0,1)N(2U,0.09)N(0.5Z,1)N(3U -Z + 0.5X , 0.25)$ and fit the model $Y \sim N(w_yx X+\beta Z,\epsilon)$ in Stan and plot the posterior distributions over $w_yx$ and $\beta$. Figure (a) shows the results with no prior (equivalently an improper prior) for both $w_yx$ and $\beta$. Figure (b) shows the results when we place a Gaussian prior centered around the $w_yz$ on the distribution for $\beta$. All models were fit using Stan \citep{carpenter2016stan}}\label{fig:careful-prior-example}
\end{figure}

@@

Reinforcement learning describes the problem of an agent interacting with an enviroment, learning by observing the outcome of its actions, with the goal of maximising some reward. These problems, which also incorporate planning, are extremely difficult because the value of an action is generally not immediately clear. The state of the enviroment may evolve over time and according to previously selected actions. Actions available at future timesteps can depend on those taken in the past, and rewards may be obtained only after a long sequence of actions. This makes it extremely difficult for the agent to acurately attribute value to each chosen action along the path to obtaining a given reward.

We focus on a simpler class of problems within reinforcement learning known as multi-armed bandit problems. Multi-armed bandit problems also descibe an agent aiming to maximise some reward by interacting with an enviroment. However, the reward is observed immediately after the action is selected and the enviroment is stateless. Future rewards generated by the  enviroment in response to the agents actions do not depend on the previous actions of the agent. These assumptions are a reasonable approximation in settings where we face the same decision repeately with respect to many independent units or individuals and we have a good proxy for the outcome. In this chapter we describe the interventionalist approaches to learning to act for these types of problems, beginning with classical randomised experiments and then the, more adaptive, multi-armed bandit algorithms. We show how both observational causal inference and bandit algorithms can be viewed as generalisations of randomized experiments and highlight connections between the observational and interventional approaches to causal inference.
